---
title: "Known markers for cell-type identification"
author: "Evelin Aasna"
date: "7/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Seurat)
load(file = "/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/integrated_seu.RData")
```


Ideas for violin plotting??
1. visualize cluster marker genes
2. visualize highly differentially expressed genes
3. visualize known markers for specific cell types

# Top 10 variable features
```{r fig.width=10, fig.height=25}
VlnPlot(seu, features = c("PAEP", "CCL20", "CXCL8", "HSPA6", "MMP3", "SPP1", "MMP1", "MMP10", "IGFBP1", "G0S2" ), 
    pt.size = 0.2, ncol = 2)
```

```{r fig.width=10, fig.height=25}
FeaturePlot(seu, reduction = "umap", features = c("PAEP", "CCL20", "CXCL8", "HSPA6", "MMP3", "SPP1", "MMP1", "MMP10", "IGFBP1", "G0S2" ), 
    pt.size = 0.2, ncol = 2)
```

# Finding cluster biomarkers i.e representative genes
Which genes separate cluster 0 from cluster 3?
```{r}
FindMarkers(seu, ident.1 = 0, ident.2 = 3, min.pct=0.25, only.pos = TRUE)[1:10,]
```


# Known markers from cell atlas of human uterus
High levels of gene expression
Endometrial epithelial: KRT8, KRT18, EPCAM and CLDN3
Endometrial stromal: MME, FN1, COL3A1, HOXA10
Endothelial: CD34, VWF, CLDN5, SOX18
SMA+(smooth muscle and myofibrolblasts): ACTA2, MYH11, MYL6, MYL9
Endometrial immune: PTPRC, CD68, CD163 and CD96

```{r}
epi_ref=c("KRT8", "KRT18", "EPCAM", "CLDN3")
stro_ref = c("MME", "FN1", "COL3A1", "HOXA10")
immu_ref = c("PTPRC","CD68", "CD136", "CD96")
```

# Known markers from molecular and cellular cartography
Decidual glandular epithelial: EPCAM, PAEP
Endothelial: CD34, CD31 antigen (PECAM1)
Heterogeneous population of fibroblasts: express matrix protein genes (e.g. PCOLCE, COL6A2) and is negative for PAEP, CD45, and CD31

```{r}
c("EPCAM", "PAEP")
c("CD34", "CD31")
c("PCOLCE", "COL6A2")
```

# Known markers from hormone responsive organoid cultures
Markers for putative stem cells: SSEA-1+(gene FUT4), SOX9+ 
```{r}
c("FUT4", "SOX9")
```


Visualize on embedding 
```{r}
FeaturePlot(seu, reduction = "umap", features = goi$endothelial, pt.size = 0.2, ncol = 3)
#ggsave("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/epi_umap_RNA.pdf", height = 10, width = 10)
```

Curated list of known markers from Angela
```{r}
goi = list( cilia = c("TUBA1A", "CTH", "EZR","FOXJ1"),

secretory = c("PAEP", "MTHFD1","ESR1"),

up_glands_vs_stroma=c("CDH1","CLDN10","EPCAM","PAX8","MUC1","PAEP",

"KLC11","MUC20","FOXA2","SOX17","KLF5"),

mucosal_secretory=c("PAX8","MUC1"),

glandular_products=c("PAEP","KLC11","MUC20"),

up_stroma_vs_gland=c("THY1","NT5E","IFITM1","COL8A1","COL12A1","COL13A1",

"LAMA1","MMP11","MMP2","MMP12","MMP27","MMP3","TIMP2","CTGF"),

endothelial=c("PECAM1","CD34","VWF"), # PECAM1 and CD34 confirmed by Roser

smooth_muscle = c("MGP"),

epithelial = c("MUC16"), # AKA CA125, in epithelia of the fallopian tube, endometrium, endocervix and ovaries

endometrial_epithelium = c("ALCAM"),

vaginal_epithelium = c("KRT4","KRT14"), # KRT4 - Predominantly non-keratinised squamous epithelia i ectocervix and vagina, KR14 is expressed in vaginal and cervical but not in uterine epithelium

stromal=c("ANPEP","VIM"),

krjutskov_endometrial_epithelial = "CD9",

krjutskov_endometrial_stromal = "ANPEP",

krjutskov_basal_glandular_epithelial_cells = "FUT4", #SSEA-1 or FUT4

petra_stem=c("MYC","CCND2","VGLL4","LCP1","LGR5","VDR","WWP1"),

epithelial_stem=c("PROM1","AXIN2","LRIG1","SOX9"),

endometrial_stem=c("FUT4","FUT9","SOX9"),

epithelial_glandular_vs_luminal_marker_glandular = c("ITGA1"),

epithelial_glandular_vs_luminal_marker_luminal = c("WNT7A","SVIL"),

epithelial_glandular_vs_luminal_top_de = c("HPGD","SULT1E1","LGR5","VTCN1","ITGA1"),

cemsel_stem=c("PDGFRB"),

gargett_mesenchymal_stem=c("ITGB1","CD44","NT5E","THY1","ENG","PDGFRB","MCAM"),

gargett_menstrual_blood_ercs=c("TERT","POU5F1","SSEA4","NANOG"), # MSCs or stromal fibroblasts, STRO-1 has no gene symbol because no seq is available

proliferation=c("MKI67"), # not expressed

differentiation=c("PAEP","SPP1", "17HSDB2", "LIF"),

stromal_decidualisation = c("IGFBP1","PRL","PRLR","MMP3","MMP9","LEFTY2","FOXO1","IL15"), # PRLR is mainly expressed in epithelial cells (IRF1)

epithelial_decidualisation = c("PRLR","IRF1","FOXO1"),

leukocytes=c("PTPRC","HLA-B"),

tcells=c("ITGAE","CD69","CXCR4","CD8A", "CD8B", "CD4"),

apoptosis=c("CASP3"),

ovary_specific=c("MUM1L1","RBP1","CDH11","NCAM1"),

endometrium_specific=c("MME"),

up_fallopian_vs_endometrium=c("FMO3"),

up_endometrium_vs_fallopian=c("DMBT1"),

cervix_fallopian_specfic=c("MUC4"),

blending=c("PTPRC_HLA-B"),

stromal_odd=c("NEAT1","MALAT1"),

msc=c("NT5E","THY1","ENG"), # and do not express CD45, CD34, ITGAX, CD14, CD19, CD79A and HLA-DR

haem=c("PTPRC","CD34","ITGAX","CD14","CD19","CD79A","HLA-DRA","HLA-DRB1"),

cycling=c("MCM2","CENPF","MK167"),  # Roser

all_mf=c("DKK1","MCAM", "PDGFRB", "REN", "RGS5","ACTA2"), # MCAM (CD146) and PDGFRB are pericyte markers

all_nk = c("ITGA1","CD9","CD39","CDHR1","CXCR4", "NCAM1", "ITGAE"), # endometrial resident NK cells

roser_pericytes = c("MCAM","PDGFRB","REN","RGS5"),

roser_glands_active = c("FXYD2","MT1G","S100A1","MT1F"),

roser_glands_exhausted = c("TMEM49","XIST","CP","VMP1","NEAT1"),

roser_decidual_stromal = c("DKK1"),

roser_spongiosa_vs_compacta = c("ACTA2"),

hla=c("HLA-A")
)

dissociation_effects=c('ACTG1','ANKRD1','ARID5A','ATF3','ATF4','BAG3','BHLHE40','BRD2',

'BTG1','BTG2','CCNL1','CCRN4L','CEBPB','CEBPD','CEBPG','CSRNP1','CXCL1','CYR61',

'DCN','DDX3X','DDX5','DES','DNAJA1','DNAJB1','DNAJB4','DUSP1','DUSP8','EGR1','EGR2',

'EIF1','EIF5','ERF','ERRFI1','FAM132B','FOS','FOSB','FOSL2','GADD45A','GCC1','GEM',

'H3F3B','HIPK3','HSP90AA1','HSP90AB1','HSPA1A','HSPA1B','HSPA5','HSPA8','HSPB1',

'HSPH1','ID3','IDI1','IER2','IER3 ','IFRD1','IL6','IRF1','IRF8','ITPKC','JUN','JUNB',

'JUND','KLF2','KLF4','KLF6','KLF9','LITAF','LMNA','MAFF','MAFK','MCL1','MIDN',

'MIR22HG','MT1','MT2','MYADM','MYC','MYD88','NCKAP5L','NCOA7','NFKBIA','NFKBIZ',

'NOP58','NPPC','NR4A1','ODC1','OSGIN1','OXNAD1','PCF11','PDE4B','PER1','PHLDA1',

'PNP','PNRC1','PPP1CC','PPP1R15A','PXDC1','RAP1B','RASSF1','RHOB','RHOH','RIPK1',

'SAT1','SBNO2','SDC4','SERPINE1','SKIL','SLC10A6','SLC38A2','SLC41A1','SOCS3','SQSTM1',

'SRF','SRSF5','SRSF7','STAT3','TAGLN2','TIPARP','TNFAIP3','TNFAIP6','TPM3','TPPP3',

'TRA2A','TRA2B','TRIB1','TUBB4B','TUBB6','UBC','USP2','WAC','ZC3H12A','ZFAND5','ZFP36',

'ZFP36L1','ZFP36L2','ZYX','GADD45G','HSPE1','IER5','KCNE4')
```

