---
title: "Known markers for cell-type identification"
author: "Evelin Aasna"
date: "7/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
```


Converting Seurat object into count matrix
```{r}
load(file = "/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_objects/joined.RData")
```

# Finding cluster biomarkers i.e representative genes
```{r}
cluster0_markers <- FindMarkers(seu, ident.1 = 0, min.pct=0.25, only.pos = TRUE)
cluster0_markers[1:10,]
```

Which genes (i.e features) separate cluster 1 from all other clusters? Only searching for positive markers that are expressed in > 25% of cells in cluster 1. 
```{r}
cluster1_markers <- FindMarkers(seu, ident.1 = 1, min.pct=0.25, only.pos = TRUE)
cluster1_markers[1:10,]
```

```{r}
cluster2_markers <- FindMarkers(seu, ident.1 = 2, min.pct=0.25, only.pos = TRUE)
cluster2_markers[1:10,]
```

```{r}
cluster3_markers <- FindMarkers(seu, ident.1 = 3, min.pct=0.25, only.pos = TRUE)
cluster3_markers[1:10,]
```

High levels of gene expression
Endometrial epithelial: KRT8, KRT18, EPCAM and CLDN3
Endometrial stromal: MME, FN1, COL3A1, HOXA10
Endothelial: CD34, VWF, CLDN5, SOX18
SMA+(smooth muscle and myofibrolblasts): ACTA2, MYH11, MYL6, MYL9
Endometrial immune: PTPRC, CD68, CD163 and CD96
```{r}
epi_ref=c("KRT8", "KRT18", "EPCAM", "CLDN3")
stro_ref = c("MME", "FN1", "COL3A1", "HOXA10")
the_ref = c("CD34", "VWF", "CLDN5", "SOX18")
muscl_ref = c("ACTA2", "MYH11", "MYL6", "MYL9")
immu_ref = c("PTPRC","CD68", "CD136", "CD96")


check_for_markers <- function(ref){
  if (length(ref[ref %in% row.names(cluster0_markers)])!=0) print(paste(0, ref[ref %in% row.names(cluster0_markers)]))
  if (length(ref[ref %in% row.names(cluster1_markers)])!=0) print(paste(1, ref[ref %in% row.names(cluster1_markers)]))
  if (length(ref[ref %in% row.names(cluster2_markers)])!=0) print(paste(2, ref[ref %in% row.names(cluster2_markers)]))
  if (length(ref[ref %in% row.names(cluster3_markers)])!=0) print(paste(3, ref[ref %in% row.names(cluster3_markers)]))
}  

check_for_markers(epi_ref)
check_for_markers(stro_ref)
check_for_markers(the_ref)
check_for_markers(muscl_ref)
check_for_markers(immu_ref)
```

Decidual glandular epithelial: EPCAM, PAEP
Endothelial: CD34, CD31 antigen (PECAM1)
Heterogeneous population of fibroblasts: express matrix protein genes (e.g. PCOLCE, COL6A2) and is negative for PAEP, CD45, and CD31

```{r}
check_for_markers(c("EPCAM", "PAEP"))
check_for_markers(c("CD34", "CD31"))
check_for_markers(c("PCOLCE", "COL6A2"))
```

