---
title: "Conos SCTransformed sample 7"
author: "Evelin Aasna"
date: "8/16/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
devtools::install_github("hms-dbmi/conos")
library(conos)
library(Seurat)
library(ggplot2)
library(cowplot)
library(dplyr)
library(SingleCellExperiment)
library(SummarizedExperiment)
library(scmap)
library(pheatmap)
source("/icgc/dkfzlsdf/analysis/B210/Evelin/git-repo/seurat_utils.R")
dir_path = "/icgc/dkfzlsdf/analysis/B210/data/mf/"
suffix = "/outs/filtered_gene_bc_matrices/GRCh38/"
```

```{r}
norm = "log"
if (norm == "log"){assay = "RNA"}
if (norm == "SCT"){assay = "SCT"}
```

Created an integrated seurat object with 4 assays. Using subset of cells from sample 07
```{r include=FALSE}
load(file = paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/integrated_seu.RData"))
seu_full = seu
barcodes07 = colnames(seu[[assay]]@data[,which(seu$sample == "07")])
seu = subset(seu, cells = barcodes07)
```

```{r }
DefaultAssay(seu) <- assay

if (norm == "log"){nPCA = 20}
if (norm == "SCT"){nPCA = 30}
seu <- RunPCA(seu, features = VariableFeatures(seu) )
seu <- RunTSNE(seu, dims = 1:nPCA )
seu <- FindNeighbors(seu, dims = 1:nPCA)
seu <- FindClusters(seu, verbose = FALSE)
```

```{r include=FALSE}
seu_list = c(seu)
```

```{r include=FALSE}
load( file = paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/decidua/",norm,"_seu_5000.RData"))
```


```{r include=FALSE}
seu_list[[length(seu_list)+1]] = seu
rm(seu)
names(seu_list) = c("menstrual", "decidua")
```

Using 5000 most variably expressed decidua genes as reference. 

Each dataset is clustered separately
```{r message = FALSE, warning=FALSE}
# how many cores on HPC?
con <- Conos$new(seu_list, n.cores=4)

con$plotPanel(clustering="multilevel", use.local.clusters=T, title.size=6)
```

```{r}
con$buildGraph(k=30, k.self=5, space='PCA', ncomps=30, n.odgenes=2000, 
               matching.method='mNN', metric='angular', 
               score.component.variance=TRUE, verbose=TRUE)

con$findCommunities(method=leiden.community, resolution=1)
```

Clusters correspond between datasets
```{r}
con$plotPanel(font.size=4)
```

```{r}
cellannot <- seu_list[["decidua"]]$annotation

# propagating labels from reference
new.label.probabilities <- con$propagateLabels(labels = cellannot, verbose=T, fixed.initial.labels=T)
new.label.probabilities <- new.label.probabilities[complete.cases(new.label.probabilities),]
```

```{r, fig.width=5, fig.height=5}
#pdf(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/",norm,"_uncertainty.pdf"))
con$plotGraph(colors=(1 - apply(new.label.probabilities, 1, max)), show.legend=T, legend.title="Uncertainty", legend.pos=c(1, 0))
#dev.off()
```

```{r include=FALSE}
new.annot <- setNames(colnames(new.label.probabilities)[apply(new.label.probabilities,1,which.max)], rownames(new.label.probabilities))
#write.table(as.data.frame(new.annot), file = "/icgc/dkfzlsdf/analysis/B210/Evelin/new_annot_sct.txt", col.names = F, quote = F, row.names = T, append = F)
```

Clusters labelled based on reference
```{r}
#pdf(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/",norm,"_conos_labelled_clusters.pdf"))
con$plotPanel(groups = new.annot)
#dev.off()
```


Projecting conos predictions onto initial UMAP from Seurat
```{r include=FALSE}
new.annot = new.annot[which(substr(names(new.annot), 1, 1) != "F")]
#save(new.annot, file = paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/",norm,"_annot"))

conos = data.frame(barcode = colnames(seu_full[[assay]])) %>% 
  left_join(data.frame(barcode = names(new.annot), celltype = unname(new.annot)), by = "barcode")

seu_full$conos = conos$celltype
```


```{r fig.height=7, fig.width=10}
DimPlot(seu_full, reduction = "log.umap", group.by = "conos", cells = which(!is.na(seu_full$conos)))
```

```{r fig.height=7, fig.width=10}
DimPlot(seu_full, reduction = "log.umap", group.by = "conos", cells = which(seu_full$conos %in% c("SCT", "VCT", "EVT", "fFB1", "fFB2"))) + labs(title = "predicted fetal cells")
```


```{r fig.height=7, fig.width=10}
DimPlot(seu_full, reduction = "log.umap", group.by = "conos", cells = which(!(is.na(seu_full$conos) | seu_full$conos %in% c("SCT", "VCT", "EVT", "fFB1", "fFB2")))) + labs(title = "predicted maternal cells")
```

# Marker gene based cluster labels
```{r fig.height=5, fig.width=15}
seu_full$celltype = 0
seu_full$celltype[which(seu_full$SCT.clusters == 14 | seu_full$log.clusters == 15) ]="endothelial"
seu_full$celltype[which(seu_full$SCT.clusters %in% c(0, 2, 3, 6, 9, 12)) ]="epithelial"
seu_full$celltype[which(seu_full$SCT.clusters %in% c(1, 5)) ]="stromal"
seu_full$celltype[which(seu_full$SCT.clusters %in% c(11)) ]="smooth muscle"
seu_full$celltype[which(seu_full$celltype==0) ]=NA

plot_grid(DimPlot(seu_full, reduction = "log.umap", group.by = "celltype", cells = which(!is.na(seu_full$celltype))), DimPlot(seu_full, reduction = "SCT.umap", group.by = "celltype", cells = which(!is.na(seu_full$celltype))), DimPlot(seu_full, reduction = "integrated.umap", group.by = "celltype", cells = which(!is.na(seu_full$celltype))), ncol = 3)
```


```{r fig.width=10, fig.height=7}
DimPlot(seu_full, reduction = "log.umap", group.by = "celltype", cells = which(!is.na(seu_full$celltype)))
```

# Scmap
Scmap uses SingleCellExperiment objects as input
```{r}
seu = seu_list[["menstrual"]]
rowData = data.frame(feature_symbol = row.names(seu[[assay]]))
colData = data.frame(Barcode = colnames(seu[[assay]]))
cluster_info = data.frame(cluster = Idents(seu))
cluster_info$Barcode = row.names(cluster_info)
colData = left_join(colData, cluster_info, by = "Barcode")


assays = return_assay(seu, assay)

se = SummarizedExperiment(assays = list(counts = assays[[1]], logcounts = assays[[2]]), rowData = rowData, colData = colData)

rm(seu)
menstrual = as(se, "SingleCellExperiment")
```

```{r}
load(file = paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/sce_RData/",norm,"_decidua_5000.RData"))
```

Feature selection
```{r}
decidua <- selectFeatures(decidua, suppress_plot = FALSE, 1000)
decidua <- indexCluster(decidua, cluster_col = "annotation")
```

Variable genes
```{r}
#pdf(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/",assay,"50_variable_decidua.pdf"))
pheatmap(metadata(decidua)$scmap_cluster_index, show_rownames = FALSE)
#dev.off()
```


```{r include=FALSE}
scmapCluster_results <- scmapCluster(
  projection = menstrual, 
  index_list = list(
    decidua = metadata(decidua)$scmap_cluster_index
  )
)
```

Celltypes in reference dataset
```{r}
df = as.data.frame(table(colData(decidua)$annotation))
ggplot(df, aes(x = reorder(Var1, -Freq), y = Freq)) + geom_col() + 
  theme_bw() + theme(axis.text.x = element_text(angle = 45)) + 
  labs(title = "decidua reference celltypes")
#ggsave(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/",assay,"_reference_celltypes.pdf"))
```

```{r}
df = as.data.frame(table(as.data.frame(scmapCluster_results)$decidua))
ggplot(df, aes(x = reorder(Var1, -Freq), y = Freq)) + geom_col() + 
  theme_bw() + theme(axis.text.x = element_text(angle = 45)) + 
  labs(title = "projected celltypes")
#ggsave(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/",assay,"_celltypes_all_genes.pdf"))
```
