---
title: "Comparing embeddings"
author: "Evelin Aasna"
date: "8/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
library(cowplot)
source("/icgc/dkfzlsdf/analysis/B210/Evelin/git-repo/marker_library.R")
```


Read in all seu object in a way that does not delete the previous one
```{r}
return_seurat_object <- function(type){
  load(file = paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/",type,"_seu.RData"))
  return(seu)  
}

integrated = return_seurat_object("integrated")
```

# Which cells express expression level for all marker genes is > n 
```{r fig.width=10, fig.height=5}
plot_grid(DimPlot(integrated, reduction = "log.umap", group.by = "log.clusters"), DimPlot(integrated, reduction = "log.umap", group.by = "SCT.clusters"), ncol = 2)
```

```{r fig.width=10, fig.height=5}
plot_grid(DimPlot(integrated, reduction = "log.umap", group.by = "SCT.clusters"), DimPlot(integrated, reduction = "log.umap", group.by = "integrated.clusters"), ncol = 2)
```

# Endothelial cells
```{r fig.width=15, fig.height=3}
DefaultAssay(integrated) <- "SCT"
integrated <- FindNeighbors(integrated, dims = 1:20, reduction = "SCT.pca", assay = "SCT")
integrated <- FindClusters(integrated, verbose = FALSE)
VlnPlot(object = integrated, features = marker.list$all_endothelial , ncol = 5)
```

TODO: use grey for unassigned celltypes or try plotting only endothelial cells since there are so few of them.
```{r fig.width=15, fig.height=5}
integrated$celltype = 0
integrated$celltype[which(integrated$SCT.clusters == 14 | integrated$log.clusters == 15) ]="endothelial"

plot_grid(DimPlot(integrated, reduction = "log.umap", group.by = "celltype"), DimPlot(integrated, reduction = "SCT.umap", group.by = "celltype"), DimPlot(integrated, reduction = "integrated.umap", group.by = "celltype"), ncol = 3)
```

```{r}
DimPlot(integrated, reduction = "integrated.umap", group.by = "celltype")
```

```{r fig.width=15, fig.height=3}
DefaultAssay(integrated) <- "RNA"
integrated <- FindNeighbors(integrated, dims = 1:20, reduction = "log.pca", assay = "RNA")
integrated <- FindClusters(integrated, verbose = FALSE)
VlnPlot(object = integrated, features = marker.list$all_endothelial , ncol = 5)
```

# Epithelial genes

```{r fig.width=7, fig.height=7}
VlnPlot(object = integrated, features = marker.list$epi_ref, ncol = 2, pt.size = 0.1 )
```


0, 2, 3, 6, 9, 12 in SCT assay

```{r fig.width=15, fig.height=5}
integrated$celltype[which(integrated$SCT.clusters %in% c(0, 2, 3, 6, 9, 12)) ]="epithelial"
integrated$celltype[which(integrated$SCT.clusters %in% c(1, 5)) ]="stromal"
integrated$celltype[which(integrated$SCT.clusters %in% c(11)) ]="smooth muscle"

plot_grid(DimPlot(integrated, reduction = "log.umap", group.by = "celltype"), DimPlot(integrated, reduction = "SCT.umap", group.by = "celltype"), DimPlot(integrated, reduction = "integrated.umap", group.by = "celltype"), ncol = 3)
```


# Cell type assignments based on decidua reference
```{r}
annot = read.table("/icgc/dkfzlsdf/analysis/B210/Evelin/SCT_annotation.txt", header = FALSE)
```


# Highly variable genes

```{r fig.width=10, fig.height=25}
DefaultAssay(integrated) <- "RNA"
FeaturePlot(integrated, features = VariableFeatures(integrated)[1:10], reduction = "log.umap", ncol = 2)
```

```{r fig.width=10, fig.height=25}
DefaultAssay(integrated) <- "SCT"
FeaturePlot(integrated, features = VariableFeatures(integrated)[1:10], reduction = "log.umap", ncol = 2)
```

```{r}
VariableFeatures(integrated)
```





---------------------------------------------------------------------------------------------------------------------
Highlight cells on UMAP embedding that express > n UMI counts of certain gene


```{r}
plot_ref <- function(seu, assay, celltype, marker_genes, threshold){
  seu$celltype = 0
  seu$celltype[which(as.matrix(seu[[assay]]@data[marker_genes, ] > threshold))]=celltype


  DimPlot(seu, reduction = "umap", group.by="celltype") + 
  labs(title = paste("putative", celltype,"cells")) +
  theme(plot.title = element_text(size = 10))
  #ggsave("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/putative epithelial.pdf")
} 
```

TODO: plot putative epithelial cells from each normalization mehtod to the same umap embedding.
Add celltype slot by finding corresponding cell barcodes.
```{r}
marker_genes = c("EPCAM", "KRT18", "KRT8", "CLDN3")
celltype = "epithelial"

plot_ref(integrated, "integrated", "epithelial", marker_genes, 1)
```


```{r}
plot_ref(sct, "SCT", "epithelial", marker_genes, 0)
```

```{r}
plot_ref(log, "RNA", "epithelial", marker_genes, 0)
```

Cluster 14 in log and 13 in sct are clearly endothelial cells. 
Where are SCT cluster 13 cells on other umap embeddings?