---
title: "Scmap - map to reference"
author: "Evelin Aasna"
date: "8/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(scmap)
library(SingleCellExperiment)
```

```{r}
load("/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/sct_cca_menstrual_sce.RData")
load("/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/decidua_sce.RData")
```


# Reference pre analysis
Selecting variable features from reference dataset
```{r}
decidua <- selectFeatures(decidua, suppress_plot = FALSE, 50)
```

Finding median gene expression for each cluster
```{r}
decidua <- indexCluster(decidua, cluster_col = "annotation")
```


```{r}
library(pheatmap)
pdf("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/50_variable_decidua.pdf")
pheatmap(metadata(decidua)$scmap_cluster_index, show_rownames = FALSE)
dev.off()
```

Selected variably expressed genes do not seem to be very variable. 

```{r}
max(metadata(decidua)$scmap_cluster_index)
```

# Projecting to reference
```{r message=FALSE, warning=FALSE}
scmapCluster_results <- scmapCluster(
  projection = menstrual, 
  index_list = list(
    decidua = metadata(decidua)$scmap_cluster_index
  )
)
```

```{r}
unique(scmapCluster_results[[1]])
```

```{r}
library(ggplot2)
```

```{r}
as.data.frame(scmapCluster_results)
```


Cell types in decidual data set
```{r}
df = as.data.frame(table(colData(decidua)$annotation))
ggplot(df, aes(x = reorder(Var1, -Freq), y = Freq)) + geom_col() + theme_bw() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "decidua reference celltypes")
ggsave("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/decidua_reference_celltypes.pdf")
```


Projected cell types in menstrual data set
```{r}
df = as.data.frame(table(as.data.frame(scmapCluster_results)$decidua))
ggplot(df, aes(x = reorder(Var1, -Freq), y = Freq)) + geom_col() + theme_bw() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "celltypes based on all genes")
ggsave("/icgc/dkfzlsdf/analysis/B210/Evelin/plots/celltypes_all_genes.pdf")
```

```{r}
df[which(df$Var1=="Epi2"), ]
```


How well do Seurat clusters and scmap projections line up? (not well)
```{r}
plot(
  getSankey(
    colData(menstrual)$cluster, 
    scmapCluster_results$scmap_cluster_labs,
    plot_height = 400
  )
)
```


Do Epi2 cells in menstrual and decidual datasets variably express the same genes?
Constructing count matrices of only Epi2 cells from both datasets.
```{r}
decidua_epi2 = logcounts(decidua)[, which(colData(decidua)$annotation=="Epi2")]
row.names(decidua_epi2) = rowData(decidua)$feature_symbol
menstrual_epi2 = logcounts(menstrual)[, which(scmapCluster_results$combined_labs == "Epi2")]
row.names(menstrual_epi2) = rowData(menstrual)$feature_symbol
```

Which genes are highly expressed in both datasets?
Just finding highly expressed genes is too simple. Need to filter out genes that are highly expressed in all cells.
Error: current and replacement dimanames() differ
```{r}
row.names(logcounts(decidua)) = rowData(decidua)$feature_symbol
de_high = names( tail( sort( rowSums( logcounts(decidua) ) ), 100) )

row.names(logcounts(menstrual)) = rowData(menstrual)$feature_symbol
me_high = names( tail( sort( rowSums( logcounts(menstrual) ) ), 100) )
```

```{r}
de_Epi2_high = names( tail( sort( rowSums( decidua_epi2 ) ), 100 ) )
me_Epi2_high = names( tail( sort( rowSums( menstrual_epi2 ) ), 100 ) )
```



Human cell atlas known markers for endometrial epithelial cells
```{r}
epi_ref=c("KRT8", "KRT18", "EPCAM", "CLDN3")
intersect(epi_ref, me_var)
```

Which genes are variable expressed.
Might be useful for finding epithelial cell subsets
```{r}
de_var = names(tail(sort(apply(decidua_epi2, 1, var)), 50))
me_var = names(tail(sort(apply(menstrual_epi2, 1, var)), 50))
length(intersect(de_var, me_var))
```

Adding decidua based projection to seurat object
```{r}
seufile = "/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/joined.RData"
seu$decidua.projection = factor(scmapCluster_results$combined_labs)
save(seu, file = seufile)
```


