---
title: "Scmap - map to reference"
author: "Evelin Aasna"
date: "8/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(scmap)
library(SingleCellExperiment)
library(ggplot2)
```


```{r}
assay = "SCT"
load(paste0("/icgc/dkfzlsdf/analysis/B210/Evelin/",assay,"_cluster_results.RData"))
unique(scmapCluster_results[[1]])
```

```{r}
as.data.frame(scmapCluster_results)
```


Do Epi2 cells in menstrual and decidual datasets variably express the same genes?
Constructing count matrices of only Epi2 cells from both datasets.
```{r}
decidua_epi2 = logcounts(decidua)[, which(colData(decidua)$annotation=="Epi2")]
row.names(decidua_epi2) = rowData(decidua)$feature_symbol
menstrual_epi2 = logcounts(menstrual)[, which(scmapCluster_results$combined_labs == "Epi2")]
row.names(menstrual_epi2) = rowData(menstrual)$feature_symbol
```

Which genes are highly expressed in both datasets?
Just finding highly expressed genes is too simple. Need to filter out genes that are highly expressed in all cells.
Error: current and replacement dimanames() differ
```{r}
row.names(logcounts(decidua)) = rowData(decidua)$feature_symbol
de_high = names( tail( sort( rowSums( logcounts(decidua) ) ), 100) )

row.names(logcounts(menstrual)) = rowData(menstrual)$feature_symbol
me_high = names( tail( sort( rowSums( logcounts(menstrual) ) ), 100) )
```

```{r}
de_Epi2_high = names( tail( sort( rowSums( decidua_epi2 ) ), 100 ) )
me_Epi2_high = names( tail( sort( rowSums( menstrual_epi2 ) ), 100 ) )
```



Human cell atlas known markers for endometrial epithelial cells
```{r}
epi_ref=c("KRT8", "KRT18", "EPCAM", "CLDN3")
intersect(epi_ref, me_var)
```

Which genes are variable expressed.
Might be useful for finding epithelial cell subsets
```{r}
de_var = names(tail(sort(apply(decidua_epi2, 1, var)), 50))
me_var = names(tail(sort(apply(menstrual_epi2, 1, var)), 50))
length(intersect(de_var, me_var))
```

Adding decidua based projection to seurat object
```{r}
seufile = "/icgc/dkfzlsdf/analysis/B210/Evelin/seurat_object/joined.RData"
seu$decidua.projection = factor(scmapCluster_results$combined_labs)
save(seu, file = seufile)
```


