---
title: "Gynecological data"
author: "Evelin Aasna"
date: "7/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages("Seurat")
library(Seurat)
library(dplyr)
library(ggplot2)
```

CellRanger output:
matrix of cells(barcodes) and genes(features) on axes
UMI - unique molecular identifier
each expressed transcript fragment in each cell has separate UMI
each UMI connected to one cell one gene

```{r}
dir_path = "/icgc/dkfzlsdf/analysis/B210/data/mf/"
run = "cellranger201_count_23156_6_GRCh38/"
#run = "cellranger201_count_23156_{6, 7, 8}_GRCh38/"
#run = "cellranger201_count_24192-25608_4839STDY713158{2, 3, 4}_GRCh38/"
```

Reading old Seurat object 
```{r}
load(file = paste0(dir_path, run, "seurat.RData"))
```

# Quality control

```{r}
seu = UpdateSeuratObject(seudata)
```


```{r}
#seu = CalculateBarcodeInflections(seu)
#Tool(seu, slot="CalculateBarcodeInflections")
```


Knee plot that helps determine cut off point for low quality (empty) cells
y <- number of cells (barcode)
x <- number of expressed genes


```{r}
n_detected_genes_sorted <- sort(seu$nGene, decreasing = TRUE)
```

```{r}
tail(n_detected_genes_sorted)
```


```{r}
plot(seq.int(length(n_detected_genes_sorted)), n_detected_genes_sorted, log = "xy", type = "s", xlab = "barcodes", ylab = "detected genes")
```


Excluding cells that expressed < n genes
```{r}
n = 600
low_exp_cells = names(seu$nFeature_RNA[seu$nFeature_RNA < n])
seu = subset(seu, cells = low_exp_cells, invert = TRUE)
```

Excluding cells that contained > 20% mitochondrial genes
```{r}
high_mt_cells = names(seu$nFeature_RNA[seu$percent.mito > 0.2])
seu = subset(seu, cells = high_mt_cells, invert = TRUE) 
```

Excluding genes that were expressed in < 3 cells
```{r}
count_matrix = as.matrix(seu[["RNA"]]@counts)
count_matrix[count_matrix > 0] <- 1
```

how many cells expressed each gene?
```{r}
cells_per_gene = rowSums(count_matrix)
rm(count_matrix)
low_expression_genes = names(cells_per_gene[cells_per_gene<3]) 
low_expression_genesS
```


nFeature <- number of genes per cell
nCount <- number of transcripts per cell
```{r}
VlnPlot(seu, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```


```{r}
head(seu$nGene)
```
```{r}
head(seu$nFeature_RNA)
```

```{r}
head(seu$nUMI)
```

```{r}
head(seu$nCount_RNA)
```


# Normalizing data
```{r}
seu <- NormalizeData(seu)
```

# Identifying highly variable features

subset of features that exbitis high cell-to-cell variation
```{r}
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(seu), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seu)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)

plot2
```
#Scaling data
```{r}
all_genes = rownames(seu)
seu <- ScaleData(seu, features = all_genes)
```

# Linear dimensionality reduction
```{r}
seu <- RunPCA(seu, features = VariableFeatures(seu))
```

```{r}
VizDimLoadings(seu, dims = 1:2, reduction = "pca")
```

```{r}
ElbowPlot(seu)
```

# Clustering

```{r}
seu <- FindNeighbors(seu, dims = 1:10)
seu <- FindClusters(seu, resolution = 0.5)
```

Each cell was assigned a cluster identity
```{r}
head(Idents(seu), 5)
```


# Non-linear dimensionality reduction
```{r}
seu <- RunUMAP(seu, dims = 1:10)
```

